---
title: "Project"
output: html_notebook
---


```{r}
install.packages("dplyr")
install.packages("Hmisc")
install.packages("ggplot2")
install.packages("tidyr")
install.packages("astsa")
install.packages("smooth")
library(cowplot)
library(imputeTS)
library(astsa)
library(ggplot2)
library(tidyr)
library(Hmisc)
library(dplyr)
```

## Reading data into R and Renaming the Variables for easier analysis.

```{r}
Climate_data <- read.csv("Kibabii Climate Data 2.csv")
Climate_data <- Climate_data %>% rename(Atmospheric_Pressure = atmosphericpressure..kPa., Precipitation = precipitation..mm., radiation = radiation..W.m2., Relative_humidity = relativehumidity...., Temperature = temperature..degrees.Celsius., Winddirection = winddirection..degrees., windgusts = windgusts..m.s., windspeed = windspeed..m.s.)
```

## Removing Unnecessary Columns

```{r}
Climate_data2 <- Climate_data %>% select(-c(2,3,4))
str(Climate_data2)
summary(Climate_data2)
```
## Removing Missing Values 

```{}
NAs_Rel_humidity <-Climate_data2 %>% subset(is.na(Relative_humidity))
NAs_Temp <-Climate_data2 %>% subset(is.na(Temperature))
NAs_winddirection <-Climate_data2 %>% subset(is.na(Winddirection))
NAs_windgusts <-Climate_data2 %>% subset(is.na(windgusts))
NAs_windspeed <-Climate_data2 %>% subset(is.na(windspeed))

```

## Imputing the missing values with the median

```{}
Climate_data2$Relative_humidity[is.na(Climate_data2$Relative_humidity)] <- median(Climate_data2$Relative_humidity, na.rm = FALSE)

```

## Visualize the missing observations

```{r}
rel_humidityts <- ts(Climate_data2$Relative_humidity)
temperaturets <- ts(Climate_data2$Temperature)
winddirectionts <- ts(Climate_data2$Winddirection)
windspeedts <- ts(Climate_data2$windspeed)
windgustts <- ts(Climate_data2$windgusts)

p1 <- ggplot_na_distribution(rel_humidityts)
p2 <-ggplot_na_distribution(temperaturets)
p3 <-ggplot_na_distribution(winddirectionts)
p4 <-ggplot_na_distribution(windgustts)
p5 <-ggplot_na_distribution(windspeedts)

plot_grid(p1,p2,p3,p4,p5, labels = c("Relative_Humidity","Temperature","Winddirection","Windgusts","Windspeed"))
```

## Interpolation of the missing values

```{r}

B1 <- plot(na_interpolation(rel_humidityts, option = "spline"))
B2 <- plot(na_interpolation(temperaturets, option = "spline"))
B3 <- plot(na_interpolation(winddirectionts, option = "spline"))
B4 <- plot(na_interpolation(windgustts, option = "spline"))
B5 <- plot(na_interpolation(windspeedts, option = "spline"))

plot_grid(B1,B2,B3,B4,B5, labels = c("Relative_Humidity","Temperature","Winddirection","Windgusts","Windspeed"))
```

## Cleaning the time series plot to eliminate outliers

```{r}
Relts <- na_interpolation(rel_humidityts, option = "spline")
tempts <- na_interpolation(temperaturets, option = "spline")
wspeedts <- na_interpolation(windspeedts, option = "spline")
wguststs <- na_interpolation(windgustts, option = "spline")
wdirects <- na_interpolation(winddirectionts, option = "spline")


Relts <- tsclean(Relts)
tempts <- tsclean(tempts)
wspeedts <- tsclean(wspeedts)
wguststs <- tsclean(wspeedts)
wdirects <- tsclean(wdirects)

```
## Augmented Dickey Fuller Test

```{r}
adf.test(Relts)
adf.test(tempts)
adf.test(wspeedts)
adf.test(wguststs)
adf.test(wspeedts)

```
## Data Splitting into training and test set
```{r}
Reltstrain <- head(Relts, round(length(Relts) * 0.9))
h <- length(Relts) - length(Reltstrain)
test <- tail(Relts, h)
plot(Reltstrain)
acf2(Reltstrain)

```

```{r}
temptstrain <- head(tempts, round(length(Relts) * 0.9))
h1 <- length(tempts) - length(temptstrain)
temptstest <- tail(tempts, h1)
plot(temptstrain)
```

```{r}
wspeedtrain <- head(wspeedts, round(length(wspeedts) * 0.9))
h2 <- length(wspeedts) - length(wspeedtrain)
wspeedtest <- tail(wspeedtrain, h2)
plot(wspeedtrain)
```


```{r}
wsgusttrain <- head(wguststs, round(length(wguststs) * 0.9))
h3 <- length(wguststs) - length(wsgusttrain)
wsguststest <- tail(wguststs, h3)
plot(wsgusttrain)
```

```{r}
wsdiretrain <- head(wdirects, round(length(wdirects) * 0.9))
h4 <- length(wdirects) - length(wsdiretrain)
wdirecttest <- tail(wdirects, h4)
plot(wsdiretrain)
```
```{r}
acf2(Reltstrain, 200)
acf2(temptstrain,200)
acf2(wspeedtrain,200)
acf2(wsgusttrain,200)
acf2(wsdiretrain,200)
```


